---
title: Automated vehicle Services for People with disabilities – Involved Responsive
  Engineering (ASPIRE Center)
author: "Sanmi"
date: "5/9/2022"
output:
  html_document:
    toc: yes
    fig_width: 10
  pdf_document:
    toc: yes
---
<a name="top"></a>
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



**Aim 3: Data synthesis, extrapolation, analysis and modeling:**  



**Questions:**  
**What are the current transportation practices of individuals with disabilities?**  

  - What are the **primary modes of transportation** (i.e. fixed route public bus, shared, personal vehicle-passenger, personal vehicle-driver, Rideshare app (Uber or Lyft), automated vehicles, etc.)?  

  - What are the **usage rates** for each mode of transportation?  

**What are the current perceptions (e.g. barriers/facilitators) of individuals with disabilities regarding accessibility of each mode of transportation?**  

  - How often to people experience barriers with getting to where they need to go in the community for these modes of transportation?
  If the traveler with a disability is a parent, child, sibling, or spouse, how does that role affect accessible travel? If the travel companion of the person with a disability is a parent, child, sibling, spouses or relative how does that role affect accessible travel?
  
  - How much do barriers with current modes of transportation and transportation related services (e.g. access to adaptive driving training programs, vehicle modifications, paratransit, ride share) affect access to accessible transportation?


**How much do barriers with transportation and transportation related services (e.g. access to adaptive driving training programs, vehicle modifications, paratransit, ride share) impact community participation?** 


  - How frequently do barriers with transportation and related services impact community participation for essential (e.g. medical, work, school, grocery etc.) and less essential (e.g. social and recreational events, etc.) activities and locations
ASPIRE Center  

  - What are the primary reasons for these problems (e.g. too costly, scheduling inconveniences, unreliable, unsafe, requires specialized equipment/adaptations, unable to accommodate additional travelers (spouse, child, service animal, attendant)  
  
  - What are the anticipated barriers/facilitators to using automated vehicles for transportation?  
  
  - What vehicle features or services would be needed to make automated vehicles accessible?  
  
  - What potential impact (influence) could accessible automated vehicles have on community participation?  
  
  
  
Others

Are there any floor/ceiling effects? If so, do the questions need to be modified?  
What is the consensus on AV/ATS amongst disability groups?  
Is there a disparity across geographical locations?  
Are there differing viewpoints across disability groups?  
What areas of autonomous vehicle development should be a priority (e.g. wheelchair securement)?  

Function to estimate fisher exact test and load packages
```{r echo=TRUE, message=FALSE, warning=FALSE}
rm(list = ls())
pack<-function(x){
  for (i in x){
    if (!require(i,character.only = T)){
      install.packages(i,dependencies = T,repos="https://cran.rstudio.com")
      require(i,character.only = T)
    }
  }
}

pvalue <- function(x, ...) {                                                                                                                                                                                                                                                                                
    # Construct vectors of data y, and groups (strata) g
    y <- unlist(x)
    g <- factor(rep(1:length(x), times=sapply(x, length)))
    #if (is.numeric(y)) {
    #    # For numeric variables, perform a standard 2-sample t-test
    #    p <- t.test(y ~ g)$p.value
    #} else 
      {
        # For categorical variables, perform a chi-squared test of independence
        p <- fisher.test(table(y, g),simulate.p.value = T)$p.value
    }
    # Format the p-value, using an HTML entity for the less-than sign.
    # The initial empty string places the output on the line below the variable label.
    c("", sub("<", "&lt;", format.pval(p, digits=3, eps=0.001)))
}

pack(c("readr","dplyr","tidyr","ggplot2", "lubridate","gtsummary","stringr","readxl","tidyr","kableExtra","table1","broom"))
pack("table1")
pack("DT")

```

Load files and create data/labels dictionary
```{r message=FALSE, warning=FALSE}
dataf <- read_excel("Clean_VoC_April_DATA.xlsx")
data_labels <- read_excel("Cleaned_VoC_April_DATA_LABELS.xlsx")
state_census<- read_csv("state census - Sheet1.csv")

labs<-colnames(data_labels)
colnames(data_labels)<-colnames(dataf)
dataf<-data_labels
dataf<-dataf[order(dataf$fwat_id),]


data.dict<-data.frame(labels=colnames(dataf),surveyq=labs)

dataf%>%
  select(dpct)%>%
  table()

#dataf<-dataf%>%
#  filter(dpct=="Person with a disability")

```


# Data dictionary
```{r}
datatable(data.dict)
```

Count conditions
```{r}
#reported conditions
```


Classify states into regions
```{r include=FALSE}
#
Northeast<-c("Connecticut",
"Maine", 
"Massachusetts",
"New Hampshire",
"Rhode Island",
"Vermont", 	
"New Jersey",
"New York",
"Pennsylvania")

Midwest<-c("Indiana",
"Illinois",
"Michigan",
"Ohio",
"Wisconsin",
"Iowa",
"Nebraska",
"Kansas",
"North Dakota",
"Minnesota",
"South Dakota",
"Missouri")

South<-c("Arkansas",
"Louisiana",
"Oklahoma",
"Texas",
"Alabama",
"Kentucky",
"Mississippi",
"Tennessee",
"Delaware",
"District of Columbia",
"Florida",
"Georgia",
"Maryland",
"North Carolina",
"South Carolina",
"Virginia",
"West Virginia")


West<-c("Alaska",
"California",
"Hawaii",
"Oregon",
"Washington",

"Montana",
"Utah",
"Nevada",
"Wyoming",
"Arizona",
"Colorado",
"Idaho",
"New Mexico")

```



```{r}
#loop to replace the participant entered state with regions and divisions
dataf$regions<-NA
dataf$divisions<-NA
for (i in 1:nrow(dataf)){
  for (j in 1:nrow(state_census)){
    if((dataf$state[i] == state_census$`State Code`[j])&
       !is.na(dataf$state[i])){
      dataf$regions[i]<-state_census$Region[j]
      dataf$divisions[i]<-state_census$Division[j]
    }
    else if (is.na(dataf$state[i])){
      dataf$regions[i]<-NA
      dataf$divisions[i]<-NA}
  }
}

#confirm replacement
dataf%>%
  select(state,regions,divisions)%>%
  datatable()

```

Disability types

```{r}

disab_type<-c("Hearing difficulty",
              "Vision difficulty",
              "Cognitive difficulty",
              "Ambulatory difficulty",
              "Self-care difficulty",
              "Independent living difficulty")
discol<-c(5:10)

for (i in 1:nrow(dataf)){
  for (j in discol){
    if (dataf[i,j]=="Checked"){dataf$disab_type[i]<-disab_type[j-4]}
    }
  }

```

<a href="#top">----------------------------------------------------------------------------------Back to top of page</a>  



```{r}

dataf$Diagnosis<-NA
dataf$Num_conditions<-NA

for (i in 1:nrow(dataf)){
  Dx<-c()
  for (j in discol){
    if (dataf[i,j]=="Checked"){Dx[length(Dx)+1]<-disab_type[j-4]}
  }
  dataf$Num_conditions[i]<-length(Dx)
  dataf$Diagnosis[i]<-paste0("",Dx,collapse = ", ")
}

dataf$Diagnosis[dataf$Diagnosis==""]<-"No checked condition"
dataf$Num_conditions<-as.character(dataf$Num_conditions)

dataf<-dataf%>%
  mutate(disab_type_2=if_else(Num_conditions>1,"Multiple Disability Types",Diagnosis))



```

## Persons with disability  

```{r}

dizz<-dataf[,c(1,4,5:10,201,202,203,11)]


dizz_na<-dizz%>%
  filter(Diagnosis=="No checked condition")
  

dizz_notna<-dizz%>%
  filter(Diagnosis!="No checked condition")


#categories of those who didnt select a condition
dizz_na%>%
  group_by(dpct)%>%
  summarize(n=n())%>%
  kable()


dataf<-dataf%>%
  filter(Diagnosis!="No checked condition")
```



```{r}

#write.csv(dizz, "dizz.csv")
#read.csv(dizz, "dizz.csv")
```

Explore raw distribution of demographic characteristics
```{r}

table1(~ ethnicity+race+gender+age+marit_stat+education+time_living_disability+regions+divisions+
         connect_transpo+comfortable_av+price_range+hous_sit+
         employment+income+veteran | disab_type_2,data = dataf,rowlabelhead = "Characteristics by disability type")

```

... Collapse

Race and ethnicity into 1
```{r}

#Race and ethnicity into 1

```

Gender into 3
```{r}

#Gender into male, female, others

```

Education into some high-school, high-school, post-high-school, advanced deg
```{r}

#Education into Education into some high-school, high-school, post-high-school, advanced deg


```


# Demographic characteristics in those with multiple conditions
```{r}

#state+
#+occupation  
table1(~ ethnicity+race+gender+age+marit_stat+education+time_living_disability+regions+divisions+
         connect_transpo+comfortable_av+price_range+hous_sit+
         employment+income+veteran | Diagnosis,data = dataf%>%filter(Num_conditions>1),rowlabelhead = "Characteristics by disability type")

```


Write function to summarize multiple choice questions
```{r}

domulti_sum<-function(df){
    datalist=list()
for (i in 1:ncol(df)){
  dat<-data.frame(
    n=table(df[,i])["Checked"],
    percent=((table(df[,i])["Checked"])/nrow(df))*100,
    row.names = colnames(df)[i]) # maybe you want to keep track of which iteration produced it?
  datalist[[i]] <- dat # add it to your list
}
    big_data = do.call(rbind, datalist)
 #   big_data<-big_data%>%
#      mutate('n (%)'=paste(n,"(",percent,")"))
    return(big_data)
  }

```

# Survey response summary  

## All participants  


### What are the current transportation practices of individuals with disabilities?  
#### Transport modes  
```{r}
#transport modes
transport_modes<-domulti_sum(dataf[,14:21])
datatable(transport_modes)
rownames(transport_modes)<-c("Licensed Driver or Permit holder of a Personally Owned or Leased Vehicle",
                        "Passenger of a Personally Owned or Leased Vehicle",
                        "Ground Public Transportation (e.g., paratransit, fixed route bus, train, subway, etc.)",
                        "Ground Private Transportation (e.g., Uber, Lyft, Taxi)",
                        "Light rail or train air taxi (flying vehicles or drones)",
                        "Air travel (e.g., planes, helicopters)",
                        "Water travel (e.g., boat, ferry, water taxi)",
                        "Others")
datatable(transport_modes)
transport_modes%>%
  mutate(x=rownames(transport_modes),
         percent=paste(round(percent,1),"%",sep = ""))%>%
  ggplot(aes(x=reorder(x,-n), y=n,label=percent))+
  geom_bar(stat="identity", color="black")+
  geom_text(nudge_y = -10)+
  xlab("Modes of transportation") + 
  ylab("Number of Participants") +  
  ggtitle(str_wrap("What are your current modes of transportation?",70))+
  scale_fill_manual(values = "#E63F00")+
  scale_x_discrete(labels = function(x) str_wrap(x, width = 18))+
  theme(
    plot.title = element_text(face="bold.italic"),
    axis.title.x = element_text(face="bold"),
    axis.title.y = element_text(face="bold"))
  



```

#### Vehicles you would be willing to use:
```{r}
#av systems
av_systems<-domulti_sum(dataf[,22:27])


rownames(av_systems)<-c("Full-size buses",
                        "Mini-buses or multi-passenger van ride-share",
                        "Van or mini-vans",
                        "Cars",
                        "Light rail or train air taxi (flying vehicles or drones)",
                        "Others")
datatable(av_systems)
av_systems%>%
  mutate(x=rownames(av_systems),
         percent=paste(round(percent,1),"%",sep = ""))%>%
  ggplot(aes(x=reorder(x,-n), y=n,label=percent))+
  geom_bar(stat="identity", color="black")+
  geom_text(nudge_y = -10)+
  xlab("Autonomous Vehicles (AV) and Autonomous Transportation Systems (ATS)") + 
  ylab("Number of Participants") +  
  ggtitle(str_wrap("If Autonomous Vehicles (AV) and Autonomous Transportation Systems (ATS) were available, please indicate the following types of vehicles you would be willing to use",70))+
  scale_fill_manual(values = "#E63F00")+
  scale_x_discrete(labels = function(x) str_wrap(x, width = 18))+
  theme(
    plot.title = element_text(face="bold.italic"),
    axis.title.x = element_text(face="bold"),
    axis.title.y = element_text(face="bold"))
  

```

#### Technologies you would be able to use to arrange a ride in an AV or ATS 
```{r}
#arrg_ride_tech
arrg_ride_tech<-domulti_sum(dataf[,29:35])
datatable(arrg_ride_tech)

```

#### Methods you would be willing to use to arrange a ride in an AV or ATS 
```{r}
#meth_arrg_ride___1
meth_arrg_ride<-domulti_sum(dataf[,37:43])
datatable(meth_arrg_ride)
```

### Activities you would use AV or ATS to access
```{r}
#activities_access___1
activities_access<-domulti_sum(dataf[,45:51])
datatable(activities_access)
```

### Services you would use AV and ATS for:
```{r}
services_use<-domulti_sum(dataf[,53:58])
datatable(services_use)
```

### If AV and ATS were available for fixed routes (fixed schedule) similar to public bus schedules, options you would use:
```{r}
fixed_rte_opt<-domulti_sum(dataf[,61:65])
datatable(fixed_rte_opt)
```

### If AV and ATS were available for custom routes (on-demand/not on a fixed schedule, customized pick-up and drop-off points), options you would use: 
```{r}
custom_rtes<-domulti_sum(dataf[,67:71])
datatable(custom_rtes)
```

### Accessibility features of an AV and/or ATS you would require
```{r}
access_features<-domulti_sum(dataf[,73:81])
datatable(access_features)
```

### What other forms of transportation would you like to be able to access using AV and ATS?
```{r}
other_forms_transpo<-domulti_sum(dataf[,85:92])
datatable(other_forms_transpo)

```

### Which of the following AV services can you see yourself using?
```{r}
av_services_using<-domulti_sum(dataf[,(96:102)])
datatable(av_services_using)
```

### Automated or automated driver assist features would you be interested in having in vehicle:
```{r}
pers_vehicle_asst_feat<-domulti_sum(dataf[,104:116])
datatable(pers_vehicle_asst_feat)
```

### Where would a personally owned AV be most valuable to you:
```{r}
pers_av_value<-domulti_sum(dataf[,118:123])
datatable(pers_av_value)

```

```{r}



```




## Caregivers of persons with disability    


----------------------------------------------------------------------

# Issues:

1. For questions "Please choose the answer which best describes you for survey
purposes." - you can only select one choice of   
 - Person with a disability  
 - Partner/Caregiver of person(s) with a disability  
 - Older Adult (65+)  
 
If a participant has all 3 roles, they will have to repeat the survey 3ce.  
This potentially introduces bias as the responses they provide at each of these instances might be different.  

They may give a question more or less thought when re-answering them.  
This could affect the internal validity of the results since responses from people who took the survey multiple times are not exactly comparable to those who took only once.  

Suggested solution: Ability to select more than one role.  



2. Related to Issue 1, dpct sums up to 585. the question is only asked if conset==yes  which isnt in all cases  

 



```{r}

#END


```



# Next Steps  

1. Chi-sq/Fischer exact for demographic characteristics
2. Cochran Q test for multiple response questions. 
3. Hierarchical Clustering to identify characteristics associated with certain responses 
4. Multinomial logistics regression to test the effects of the identified characteristics dimension reduction


## Modeling.

A1. Feature selection using Principal Component Analysis  
2. Identify latent structures related to geography  

3a. Evaluate features based on Eigenvalues and factor matrix (retain those >1)  

3b. Evaluate features with permutation tests to test significance

B. Neural network node with RBF to identify dimensionality of the variables.  

C. XGBoost 

5. Perform Akaike Information Criteria (AIC) for each model  


6. Cluster analysis using KMC




```{r}

#transport modes cochran Q test
dataf_tx<-dataf[,c(1,14:20)]
for (i in 1:nrow(dataf_tx)){
  for (j in 1:ncol(dataf_tx)){
      if (dataf_tx[i,j]=="Checked"){dataf_tx[i,j]<-"1"}
    if (dataf_tx[i,j]=="Unchecked"){dataf_tx[i,j]<-"0"}
  }
}

```

